Investigating Determinism of Gaming Engines for Autonomous Vehicle Verification

### 2020-02-07 paper/3_Methodology.tex
update to para 1-3

In this section a method for determining the \textit{simulation variance} is presented along with experimental parameters that may influence it, see Fig.~\ref{method_diagram}. The method includes scenarios that exercise different actor types and functional callbacks of the game engine that may influence thread scheduling. To replicate high computational loads, and hence game loop update rates, that may be experienced in high realism simulations, resource utilisation software is used at runtime for both central and graphical processing units.

\subsection{Experimental Aims}
The aim of the experiment is to determine if the simulation is deterministic. This is achieved by monitoring actors in repeated tests and examining the deviation in path trajectories. Actors are observed in a number of different scenarios involving a mixture of vehicles and pedestrians exercising the demands outlined in Section~\ref{s:background}. The user must specify the level of variance sufficient for their application which is used as a threshold for deterministic execution. In our case study, Section~\ref{s:case-study}, a threshold of $1cm$ was chosen for our given application which reflects the scale at which repeatability of simulation execution is required. This value is sufficiently small to allow accurate \textit{assertion checking}[ref] in urban environments but a larger value may be acceptable for non-urban environments, e.g. motorway/highway. Similarly, the threshold value should not be so small as to fail with minor computational bottlenecks, e.g. $1\mu m$.

@@

Tests should be sufficiently repeated to ensure that statistical power~\cite{cohen2013statistical} reveals any outlier errors that occur with low probability. Initial testing with the Unreal Engine indicated an actor path deviation of $1\times10^{-13} cm$ for 998 out of 1000 tests but with 3 tests reporting a deviation of over {\raise.17ex\hbox{$\scriptstyle\sim$}}$10cm$. Executing 100 repeats may seem sufficient but in this case would fail to observe these events that occur with low probability, inferring a false confidence in the results.

>>>>

In this section a method for determining the \textit{simulation variance} is presented along with experimental parameters that may influence it, see Fig.~\ref{method_diagram}. The method includes scenarios that exercise different actor types and functional callbacks of the game engine that may tax different aspects of the engine such as actor navigation and collision callbacks. To replicate high computational loads that may be experienced in high realism simulations, resource utilisation software is used at runtime for both central and graphical processing units. The following section outlines a guide that may be used to determine the \textit{simulation variance} of an end user system.

The aim of the experiment is to determine if the simulation is deterministic. This is achieved by monitoring actors in repeated tests and examining the deviation in path trajectories. Actors are observed in a number of different scenarios involving a mixture of vehicles and pedestrians exercising the demands outlined in Section~\ref{s:background}.

\subsection{Deterministic Threshold}
We use the term \textit{deterministic threshold} to refer to the acceptable level of actor path variance ($sigma^2$), which is quoted in terms of deviation ($sigma$) for ease of interpretation. This threshold is analogous to the analogue level associated with a binary state in a logical processor. At some low analogue voltage the signal is interpreted as binary $0$ which may be below 0.5V and is considered binary $1$ when the same signal is above 3.5v.

The user must specify the level of variance sufficient for their application which is used as a threshold for deterministic execution. In our case study, Section~\ref{s:case-study}, a threshold of $1cm$ was chosen for our given application which reflects the scale at which repeatability of simulation execution is required. This value is sufficiently small to allow accurate \textit{assertion checking} of autonomous vehicles in urban environments [ref] but a larger value may be acceptable for non-urban environments, e.g. motorway/highway. Similarly, the threshold value should not be so small as to fail with minor computational bottlenecks, e.g. $1\mu m$.

@@

Tests should be sufficiently repeated using statistical power~\cite{cohen2013statistical} to reveal any outlier errors that occur with low probability. Initial testing\footnote{Experiments using the Unreal Engine not reported here for brevity but code and experimental results can be found at \url{https://github.com/TSL-UOB/CAV-Determinism/tree/master/UnrealEngineTests}} indicated an actor path deviation of $1\times10^{-13} cm$ for 997 out of 1000 tests but with 3 tests reporting a deviation of over {\raise.17ex\hbox{$\scriptstyle\sim$}}$10cm$. While executing 100 repeats may seem sufficient, in this case the sample size would fail to observe these events that occur with low probability, inferring a false confidence in the results.
